---
title: "Score AQoL-6D (Adolescent) Health Utility with ScorzAqol6Adol"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Score AQoL-6D (Adolescent) Health Utility with ScorzAqol6Adol}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Note: **This vignette is illustrated with fake data**. The dataset explored in this example should not be used to inform decision-making. Some of the methods illustrated in this AQoL-6D vignette can also be used to score other health utility instruments - see [a vignette about scoring EQ-5D](V_02.html).

```{r warning=FALSE, message=FALSE}
library(scorz)
```

## AQoL-6D scoring
To derive a health utility score from the raw responses to a multi-attribute utility instrument it is necessary to implement a scoring algorithm. Scoring algorithms for the Assessment of Quality of Life Six Dimension (AQoL-6D) are publicly available in SPSS format (https://www.aqol.com.au/index.php/scoring-algorithms). 

However, to include scoring algorithms in reproducible research workflows, it is desirable to have these algorithms available in open science languages such as R. We therefore developed an R implementation of the adult and adolescent versions of the AQoL-6D scoring algorithms and have made them available as part of the `scorz` package.

## Ingest data
To begin, we ingest an unscored dataset as an instance of the `Ready4useDyad` class (from the [ready4use package](https://ready4-dev.github.io/ready4use/)). In this case we download our data from a remote repository.

```{r}
x <- ready4use::Ready4useRepos(dv_nm_1L_chr = "fakes",
                               dv_ds_nm_1L_chr = "https://doi.org/10.7910/DVN/W95KED",
                               dv_server_1L_chr = "dataverse.harvard.edu") %>%
  ingest(fls_to_ingest_chr = "ymh_clinical_dyad_r4",
         metadata_1L_lgl = F) 
```

To make the ingested dataset easier to interpret, we can add labels from the dictionary.

```{r}
x <- x %>%
  renew(type_1L_chr = "label")
```

We can now inspect our ingested dataset using the `exhibit` method.

```{r}
exhibit(x,
        display_1L_chr = "head")
```

We now add meta-data that identifies our dataset as being longitudinal using the `YouthvarsSeries` class of the [youthvars package](https://ready4-dev.github.io/youthvars/).

```{r}
y <- youthvars::YouthvarsSeries(a_Ready4useDyad = x,
                                id_var_nm_1L_chr = "fkClientID",
                                timepoint_var_nm_1L_chr = "round",
                                timepoint_vals_chr = levels(x@ds_tb$round))
```

We now use the data and meta-data we have created in the previous steps to create an instance of the `ScorzAqol6Adol` class. This class is specifically designed to facilitate scoring of the adolescent version of the AQoL-6D instrument.

```{r}
z <- ScorzAqol6Adol(a_YouthvarsProfile = y)
```

By default, instances of the `ScorzAqol6Adol` class are created with a slot specifying a value for the prefix for AQoL-6D questionnaire item responses.

```{r}
procureSlot(z,
            slot_nm_1L_chr = "itm_prefix_1L_chr")

```
If this default value needs to be updated to match the prefix used in your dataset, use the `renewSlot` method.

```{r}
# Not run
# z <- renewSlot(z, slot_nm_1L_chr = "itm_prefix_1L_chr", new_val_xx = "new_prefix")
```


## Calculating scores
To calculate AQoL 6D adolescent utility scores, use the `renew` method. 

```{r results='hide', warning=F, message=FALSE}
z <- renew(z)
```

## Viewing the updated dataset
We can inspect our updated dataset using the `exhibit` method. We can see that the updated dataset now has additional variables that include the intermediate and final calculations for AQoL-6D adolescent utility scores.

```{r}
exhibit(z,
        display_1L_chr = "head")
```

## Creating summary plots
To create plots, we use the `depict` method.

We can create a list of summary plots by timepoint for all individual items.

```{r}
plot_ls <- depict(z, type_1L_chr = "item_by_time")
```

We can then select a desired item's summary plot by using its index number.

```{r item1fig, fig.cap="AQoL-6D Item 1 scores by data-collection round", out.width = "600px"}
plot_ls[[1]]
```

Alternatively, we can generate individual plots by passing the item index number to the `var_idcs_int` argument of `depict`.

```{r item2fig, fig.cap="AQoL-6D Item 2 scores by data-collection round", out.width = "600px"}
depict(z, type_1L_chr = "item_by_time", var_idcs_int = 2L)
```

We can also plot domain scores by time.

```{r domain1fig, fig.cap="AQoL-6D Independet Living Domain weighted scores by data-collection round", , out.width = "600px"}
depict(z, type_1L_chr = "domain_by_time", var_idcs_int = 1L)
```

Total AQoL-6D scores can also be plotted using the same approach, where `var_idcs_int = 1L` is used to plot the weighted total distribution and `var_idcs_int = 2L` is used for plotting the unweighted total.

```{r totalsfig, fig.cap="AQoL-6D item total weighted scores by data-collection round", out.width = "600px"}
depict(z, type_1L_chr = "total_by_time", var_idcs_int = 1L)
```

Composite plots can be generated as well, though these are not currently optimised to reliably produce quality plots suitable for publication.

```{r itemsfig, fig.cap="AQoL-6D item responses by data-collection round", out.width = "600px", results='hide',fig.keep='all'}
depict(z, type_1L_chr = "comp_item_by_time")
```

```{r domainsfig, fig.cap="AQoL-6D weighted domain scores by data-collection round", out.width = "600px", results='hide', fig.keep='all'}
depict(z, type_1L_chr = "comp_domain_by_time")
```
