---
title: "Score EQ-5D Health Utility"
output: rmarkdown::html_vignette
# output: 
#   rmarkdown::html_vignette:
#     toc: true
#   pdf_document:
#     highlight: null
#     number_sections: yes
vignette: >
  %\VignetteIndexEntry{Score EQ-5D Health Utility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Note: **This vignette is illustrated with fake data**. The dataset explored in this example should not be used to inform decision-making. Some of the methods illustrated in this EQ-5D vignette are described in more detail in [a vignette about scoring AQoL-6D health utility](https://ready4-dev.github.io/scorz/articles/V_01.html). Furthermore, the synthetic data and model types used in this data have not been optimised for the purposes of this example (the models created below perform poorly as a result).

```{r warning=FALSE, message=FALSE}
library(ready4)
library(scorz)
```

## EQ-5D scoring
The [eq5d package](https://github.com/fragla/eq5d) provides a toolkit for scoring EQ-5D health utility. The  `scorz` package contains [ready4 framework model modules](https://www.ready4-dev.com/docs/framework/implementation/modularity/#ready4-model-modules) that integrate eq5d package tools with [the ready4 youth mental health economic model](https://www.ready4-dev.com/docs/model/).

## Ingest and describe data
To begin, we ingest an unscored dataset and its data dictionary as a [Ready4useDyad](https://ready4-dev.github.io/ready4use/articles/V_02.html). We then add additional details about the dataset (unique identifier and data-collection round variable name and values) by converting the ingested data to an instance of the (YouthvarsSeries module)[https://www.ready4-dev.com/docs/model/modules/using-modules/people/add-metadata/].

```{r}
X <- ready4use::Ready4useRepos(dv_nm_1L_chr = "fakes",
                               dv_ds_nm_1L_chr = "https://doi.org/10.7910/DVN/W95KED",
                               dv_server_1L_chr = "dataverse.harvard.edu") %>%
  ingest(fls_to_ingest_chr = "ymh_eq5d_dyad_r4",
         metadata_1L_lgl = F) 
X <- youthvars::YouthvarsSeries(a_Ready4useDyad = X,
                                id_var_nm_1L_chr = "uid",
                                timepoint_var_nm_1L_chr = "Timepoint",
                                timepoint_vals_chr = unique(X@ds_tb$Timepoint))
```

We now use the data and meta-data we have created in the previous steps to create an instance of the `ScorzEuroQol5` class. Important arguments to specify when creating a `ScorzEuroQol5` instance are `country_1L_chr` (the country preference weights to use), `instrument_version_1L_chr` (instrument version) and the `itm_var_nms_chr` (item variable names - **these must be in the exact order that corresponds to the Movement, Self-care, Usual Activities, Pain/Discomfort and Anxiety Depression domains**).

```{r}
Y <- ScorzEuroQol5(a_YouthvarsProfile = X,
                   country_1L_chr = "UK",
                   instrument_version_1L_chr = "5L",
                   itm_var_nms_chr = c("eq5dq_MO", "eq5dq_SC", "eq5dq_UA", "eq5dq_PD", "eq5dq_AD"))
```

## Calculating scores
To calculate EQ-5D adolescent utility scores, use the `renew` method. 

```{r results='hide', warning=F, message=FALSE}
Y <- renew(Y)
```

## Viewing the updated dataset
We can inspect our updated dataset using the `exhibit` method. 

```{r}
exhibit(Y,
        display_1L_chr = "head",
         scroll_box_args_ls = list(width = "100%"))
```

## Creating summary plots
To create plots, we use the `depict` method, supplying the index number of the dimension we wish to plot.

```{r adfig, fig.cap="EQ-5D Anxiety / Depression Dimension scores by time", out.width = "600px"}
depict(Y, type_1L_chr = "domain_by_time", var_idcs_int = 5L)
```

Total EQ-5D scores can also be plotted using the same approach, where `var_idcs_int = 1L` is used to plot the weighted total distribution.

```{r totalsfig, fig.cap="EQ-5D total weighted utility scores by time", out.width = "600px"}
depict(Y, type_1L_chr = "total_by_time", var_idcs_int = 1L)
```

A composite plot can also be generated.

```{r domainsfig, fig.cap="EQ-5D weighted domain scores by time", out.width = "600px", results='hide', fig.keep='all'}
depict(Y, type_1L_chr = "comp_domain_by_time")
```

## Share output
We can now publicly share our scored dataset and its associated metadata, using `Ready4useRepos` and its `share` method [as described in a vignette from the ready4use package](https://ready4-dev.github.io/ready4use/articles/V_01.html).

```{r eval=F}
Z <- ready4use::Ready4useRepos(gh_repo_1L_chr = "ready4-dev/scorz", # Replace with details of your repo.
                               gh_tag_1L_chr = "Documentation_0.0") # You must have write permissions.
Z <- share(Z,
           obj_to_share_xx = Y,
           fl_nm_1L_chr = "ymh_ScorzEuroQol5")
```

`Y` is now available for download as the file `ymh_ScorzEuroQol5.RDS` from the ["Documentation_0.0" release of the scorz package](https://github.com/ready4-dev/scorz/releases/tag/Documentation_0.0).
